A Survey of Near-Term Exoskeleton Technology\\ (Draft 3.0)

### 2015-09-28 exos/hal.tex
cga-fixes
Added bibliographies by section to survey.tex. Use biber to build survey.tex

\subsection{HAL}
\label{exo:hal}

@@

The hybrid assistive limb (HAL) exoskeleton is developed by the University of Tsukuba and Cyberdyne for human strength augmentation and as an assistive gait device in rehabilitation.  Although several prototypes have been developed, this section focuses on the most recent HAL-5 model (specifically the non-clicial type, HAL-5 Type-B). Though HAL-5 is a full body exoskeleton, only hip, knee joints, and ankle are actuated in the sagittal plane.  The exoskeleton powers DC motors with harmonic drives at hip, knee, ankle joints (in some models the ankle joints act as passive springs).  HAL weighs 23kg, includes an on-board AC100V battery as the power source, which is designed to support the power requirements of maximum velocity human walking and standing torque requirements.  The battery allows for 160 min of continuous operation and enables the exoskeleton to lift up to 70kg \cite{HALassist2011}.

Human operators attach to HAL at the waist by a belt, and at the calf and thigh using harness. HAL's frame does not transfer load to the ground.  Instead, HAL adjusts hip, knee and ankle torques to amplify its operator's torque. The exoskelton's sensory system includes bioelectric sensing (including sEMG), angular sensors, acceleration sensors, and center of pressure / center of gravity (COP/COG) sensors.  COP/COG sensing is provided through shoes with ground reaction force sensors.  The joint measurements are provided by potentiometers.
In at least one version of HAL the bioelectric data comes from sEMG sensors installed below the operator's hip and above the knee (on both front and back).
An IMU installed in HAL's backpack is used to estimate torso pose.

@@

HAL uses two types of control systems depending on application domain.  For gait assistance and rehabilitation, HAL uses an autonomous control system that carries the user through predefined gait trajectories by controlling knee and hip joints (the ankles behave as passive springs).  Gait phase intention is estimated from COP/COG sensors.  The exoskeleton drives the wearer to follow pre-recorded desired joint patterns.

@@

In \cite{HALmuscleImped2005}, an impedance control strategy controls the viscoelastic properties of HAL's knee joint from a musculoskeletal model of operator's limb moving in concert with the exoskeleton.  The controller uses sEMG sensor data to estimate muscle torque based on the difference between flexor and extensor muscle activities.  The authors note the sEMG model requires significant calibration effort.  
%
\begin{figure}[ht]
  \centering
  \includegraphics[width=6in]{exos/figs/hal_viscoelastic_control.png}
\end{figure}
%
%
Viscoelastic torques are computed according to
\[\tau_{a,i} = \alpha_i(-D_i \theta_i - K_i \theta_i),\]  
based on a variable gain set by $\alpha_i$.  The net knee actuator torque is
\[\tau_i = \tau_{a,i} + \tau_\mu + \tau_c .\]
The torque includes a $\tau_c$ term compensating for mechanical (actuator) impedance and $\tau_\mu$, a scaled version of the human torque (as estimated  from sEMG data).  The $K_i$ and $D_i$ terms in $\tau_{a,i}$ are viscoelastic parameters (based on the operator's muscles) in the human-exoskeleton model.  These parameters are estimated on-line using a weighted least-squares method.  In this case, the angular velocity data required for the impedance controller is estimated from a state observer.

While experiments in \cite{HALmuscleImped2005} only consider leg swing-up and swing-down, a similar control approach in \cite{HALvTorqueImp2002} switches the dynamics based controller to compensate for swing and stance phases in walking gait.  In this case, the gait transitions are detected by thresholding foot sensors and required angular velocity (and acceleration) data is determined by numerically differentiating angular encoders.

@@

Compared to sensitivity amplification methods, HAL's use of EMG data allows it to estimate intention without amplifying external disturbances.  This is ideal for combat scenarios.  However, sEMG data is extremely difficult to work with due to high filtering and calibration requirements.  A hybrid strategy which uses sensitivity amplification and sEMG data (possibly thresholded) to filter user intention could prove effective.  As an alternative, research into nano-fabrication and MEMs technologies has produced new ``robotic skin'' sensing arrays. This technology may allow direct measurement of interaction forces (i.e. estimating pressure / force and exoskeleton contact locations), which has not been feasible in the past.  Such measurements could (potentially) filter external disturbances from user generated input to estimate intent.

%\bibliographystyle{plain}
%\bibliography{exos/hal}

The figures in this section were obtained from \cite{HALassist2011,HALmuscleImped2005}.  Materials presented are based on the references above.

>>>>

\section{HAL}
\label{exo:hal}
\begin{refsection}[exos/hal.bib]

@@

The hybrid assistive limb (HAL) exoskeleton is developed by the University of Tsukuba and Cyberdyne for human strength augmentation and as an assistive gait device in rehabilitation.  Although several prototypes have been developed, this section focuses on the most recent HAL-5 model (specifically the non-clicial type, HAL-5 Type-B). Though HAL-5 is a full body exoskeleton, only hip, knee joints, and ankle are actuated in the sagittal plane.  The exoskeleton uses DC motors with harmonic drives at hip, knee, ankle joints (in some models the ankle joints act as passive springs).  HAL weighs 23kg and includes an on-board AC100V battery as the power source, which is designed to support maximum velocity human walking and standing torque requirements.  The battery allows for 160 min of continuous operation and enables the exoskeleton to lift up to 70kg \cite{HALassist2011}.

Human operators attach to HAL at the waist with a belt, and at the calf and thigh using harnesses. HAL's frame does not transfer load to the ground.  Instead, HAL adjusts hip, knee and ankle torques to amplify its operator's torque. The exoskelton's sensory system includes bioelectric sensing (including \textbf{sEMG}), \textbf{angular sensors}, \textbf{acceleration sensors}, and \textbf{center of pressure / center of gravity} (COP/COG) sensors.  COP/COG sensing is provided through shoes with ground reaction force sensors.  The joint measurements are provided by potentiometers.
In at least one version of HAL the bioelectric data comes from sEMG sensors installed below the operator's hip and above the knee (on both front and back).
An IMU installed in HAL's backpack is used to estimate torso pose.

@@

HAL has two types of control systems designed for different application domains.  For gait assistance and rehabilitation, HAL uses an autonomous control system that carries the user through predefined gait trajectories by controlling knee and hip joints (the ankles behave as passive springs).  Gait phase intention is estimated from COP/COG sensors.  The exoskeleton drives the wearer to follow pre-recorded desired joint patterns.

@@

In \cite{HALmuscleImped2005}, an impedance control strategy controls the viscoelastic properties of HAL's knee joint from a musculoskeletal model of operator's limb moving in concert with the exoskeleton.  The controller uses sEMG sensor data to estimate muscle torque based on the difference between flexor and extensor muscle activities.  The authors note the sEMG model requires significant calibration effort.  
Viscoelastic torques are computed according to
\[\tau_{a,i} = \alpha_i(-D_i \theta_i - K_i \theta_i),\]  
based on a variable gain set by $\alpha_i$.  The net knee actuator torque is
\[\tau_i = \tau_{a,i} + \tau_\mu + \tau_c .\]
The torque includes a $\tau_c$ term compensating for mechanical (actuator) impedance and $\tau_\mu$, a scaled version of the human torque (as estimated  from sEMG data).  The $K_i$ and $D_i$ terms in $\tau_{a,i}$ are viscoelastic parameters (based on the operator's muscles) in the human-exoskeleton model.  These parameters are estimated on-line using a weighted least-squares method.  In this case, the angular velocity data required for the impedance controller is estimated from a state observer.

%
\begin{figure}[ht]
  \centering
  \includegraphics[width=6in]{exos/figs/hal_viscoelastic_control.png}
\end{figure}
%
%

While experiments in \cite{HALmuscleImped2005} only consider leg swing-up and swing-down, a similar control approach in \cite{HALvTorqueImp2002} switches the dynamics based controller to compensate for swing and stance phases in walking gait.  In this case, the gait transitions are detected by thresholding foot sensors and required angular velocity (and acceleration) data are determined by numerically differentiating angular encoders.

@@

Compared to sensitivity amplification methods, HAL's use of EMG data allows it to estimate intention without amplifying external disturbances.  While direct sensing of user intention is ideal for combat scenarios, sEMG data is extremely difficult to work with due to high filtering and calibration requirements.  A hybrid strategy which uses sensitivity amplification and sEMG data (possibly thresholded) to filter user intention could prove effective.  As an alternative, research into nano-fabrication and MEMs technologies has produced new ``robotic skin'' sensing arrays. This technology may allow direct measurement of interaction forces (i.e. estimating pressure / force and exoskeleton contact locations), which has not been feasible in the past.  Such measurements could (potentially) filter external disturbances from user generated input to estimate intent.

\nocite{*}
\printbibliography[heading=subbibliography]

The figures in this section were obtained from \cite{HALmuscleImped2005,HALassist2011}.  Materials presented are based on the references above.

\end{refsection}

### 2015-09-28 exos/xor.tex
cga-fixes
Added bibliographies by section to survey.tex. Use biber to build survey.tex

\subsection{XoR}
\label{exo:XoR}

@@

The XoR is a prototype, light-weight lower-body exoskeleton that uses a hybrid pneumatic-electric drive system that \textbf{reduces weight} while providing \textbf{precise torque control}, \textbf{backdrivability}, and a \textbf{desirable force / velocity} profile.  The exoskeleton is designed to serve in rehabilitation settings to augment operators' strength and assist with postural control for persons with disabilities.

@@

The XoR weighs 30kg and includes 10 DOF with 6 active joints (flexion / extension of hip, knee, and ankles) and 6 passive (hip abduction/adduction joints, hip rotation, and ankle adduction/abduction).  The active joints are powered by hybrid actuators comprised of an \emph{air muscle} and an electric motor.  The hybrid actuation scheme uses a unilateral air muscle layout to compensate for gravity and bilateral electric motors with relatively small gear ratio (57.5) to serve as the dynamic compensator.

@@

A challenge in using air muscles is that force reduces quadratically as a function of contraction.  For instance, at 30\% contraction, the force produced by FESTO rubber air muscles vanishes.  To address the issue, XoR is strategically places air muscles to generate forces in desired configurations.

@@

In implementation, the XoR uses rubber air muscles (MDSP-40) connected to pulleys via tendons and to EP regulators in the backpack (a constant pressure and cam system is being tested to reduce weight required for backpack air valves).  The electric actuator is a geared, 200W brushless DC motor (Maxon EC powermax 30) rated at 4.7 A, which transmits power to joints through a belt and pulley system with a 57.5 gear ratio. the motor's 0.12 Nm output torque provides up to 34.5 Nm (for short duration) at every joint and is backdrivable.  The design yields a relatively high range of motion (up to 120 degrees) with a load capacity 150 Nm.  However, experiments reveal the need for bi-directional pneumatic actuation at hip joints.

For sensing, XoR is equipped with rotary encoders at joints, an IMU in the backpack, load cells in the feet, and is networked to servers capable of providing EMG, NIRS, and EEG data.  A control PC running at 1 KHz, pulse counter, amplifiers, Digital IO are external and so not included in the weight of the unit.  Human operators are outfitted with a goniometer.

@@

To avoid relying on high gain feedback, XoR incorporates user intent through EMG data.  The controller rectifies and low pass filters (10 Hz cut-off) data measured in the pilot's quadriceps femoris, tensor fasciae latate, gluteus medius, and tibialis anterior. 
It sets desired joint angles and velocities, ${\bf x} = ({\bf \theta} \dot{\bf \theta})$, from the EMG data, ${\bf u} = (EMG_1 \cdot EMG_n)$, using a linear prediction model, 
\[{\bf x}(k+1) = \mA{\bf x}(k)+\mB{\bf u}(k),\].  
The controller uses the angular measurements required to perform CCA and EMG data to derive the model's $\mA$ and $\mB$ terms.  The predicted state from EMG model provides the desired input to yet another PD controller: 
\[\tau_i = K_p (\theta_i^d - \theta_i) + K_d (\dot\theta_i^d - \dot \theta_i),\] 
with gains of $K_p = 1000$ and $K_d = 100$.

@@

The XoR prototype uses a promising hybrid actuation scheme to achieve both fast peak torque response and higher sustained torque while remaining lightweight.  They track human motion with feedback from a goniometer and feedforward control provided by EMG data.  These human intention / sensing modalities are both difficult to implement due to noise and calibration issues.  However, capturing the user intent is important, especially in the presence of external disturbances.  If the sensory system were improved (e.g, wearable ``robot skin'' sensing arrays), this exoskeleton design could be highly effective.

% \bibliographystyle{plain}
% \bibliography{exos/xor}

>>>>

\section{XoR}
\label{exo:XoR}
\begin{refsection}[exos/xor.bib]

@@

The XoR is a prototype light-weight, lower-body exoskeleton that uses a hybrid pneumatic-electric drive system that \textbf{reduces weight} while providing \textbf{precise torque control}, \textbf{backdrivability}, and a \textbf{desirable force / velocity} profile.  The exoskeleton is designed to serve in rehabilitation settings to augment operators' strength and assist with postural control for persons with disabilities.

@@

The XoR weighs 30kg and includes 10 DOF with 6 active joints (flexion / extension of hip, knee, and ankles) and 6 passive (hip abduction/adduction joints, hip rotation, and ankle adduction/abduction).  The active joints are powered by \textbf{hybrid actuators} comprised of an \textbf{air muscle} and an \textbf{electric motor}.  The hybrid actuation scheme uses a unilateral air muscle layout to compensate for gravity and bilateral electric motors with relatively small gear ratio (57.5) to serve as the dynamic compensator.

@@

A challenge in using air muscles is that force reduces quadratically as a function of contraction.  For instance, at 30\% contraction, the force produced by FESTO rubber air muscles vanishes.  To address the issue, XoR's air muscles are strategically placed to generate forces in desired configurations.

@@

In implementation, the XoR uses rubber air muscles (MDSP-40) connected to pulleys via tendons and to EP regulators in the backpack (a constant pressure and cam system is being tested to reduce weight required for backpack air valves).  The electric actuator is a geared, 200W brushless DC motor (Maxon EC powermax 30) rated at 4.7 A, which transmits power to joints through a belt and pulley system with a 57.5 gear ratio. The motor is backdrivable with 0.12 Nm output torque and up to 34.5 Nm (for short duration) at every joint.  The design yields a relatively high range of motion (up to 120 degrees) with a 150 Nm load capacity.  However, experiments reveal the need for bi-directional pneumatic actuation at hip joints.

For sensing, XoR is equipped with \textbf{rotary encoders} at joints, an \textbf{IMU} in the backpack, \textbf{load cells} in the feet, and is networked to servers capable of providing \textbf{EMG}, NIRS, and EEG data.  A control PC running at 1 KHz, pulse counter, amplifiers, Digital IO are external and so not included in the weight of the unit.  Human operators are outfitted with a \textbf{goniometer}.

@@

To avoid relying on high gain feedback, XoR incorporates user intent through EMG data.  The controller rectifies and low pass filters (10 Hz cut-off) data measured in the pilot's quadriceps femoris, tensor fasciae latate, gluteus medius, and tibialis anterior. 
It sets desired joint angles and velocities, ${\bf x} = ({\bf \theta}, \dot{\bf \theta})$, from the EMG data, ${\bf u} = (EMG_1 \dots EMG_n)$, using a linear prediction model, 
\[{\bf x}(k+1) = \mA{\bf x}(k)+\mB{\bf u}(k),\].  
The controller uses the angular measurements required to perform CCA and EMG data to derive the model's $\mA$ and $\mB$ terms.  The predicted state from EMG model provides the desired input to yet another PD controller: 
\[\tau_i = K_p (\theta_i^d - \theta_i) + K_d (\dot\theta_i^d - \dot \theta_i),\] 
with gains of $K_p = 1000$ and $K_d = 100$.

@@

The XoR prototype uses a promising hybrid actuation scheme to achieve both fast peak torque response and higher sustained torque while remaining lightweight.  It tracks human motion with feedback from a goniometer and feedforward control provided by EMG data.  These human intention / sensing modalities are both difficult to implement due to noise and calibration issues.  However, capturing the user intent is important, especially in the presence of external disturbances.  If the sensory system were improved (e.g, wearable ``robot skin'' sensing arrays), this exoskeleton design could be highly effective.

\nocite{*}
\printbibliography[heading=subbibliography]
